Fast Algorithms for Intersection of Non-matching Grids Using \plucker coordinates.

### 2016-09-30 paper_intersections.tex
Finished 23d

\subsection{Intersection Triangle-Tetrahedron (2D-3D)}
The intersection of a triangle $S_2$ and a tetrahedron $S_3$ is an $n$-side intersection polygon (IP), $n\le 7$. The sides of the polygon
lie either on sides of $S_2$ or on faces of $S_3$. Thus each vertex (IC) of the polygon
can arise either from side-face intersection, or from edge-triangle intersection, or be a vertex of $S_2$.
To get all ICs, we have to compute at most 12 side-face intersections and at most 6 edge-triangle intersections. However,
to this end we only need to compute 9 \plucker coordinates (3 sides, 6 edges) and 18 permuted inner produts, one for every side-edge pair.
Computation of IP consists of three stages: calculation of side-tetrahedron ICs (Algorithm \ref{algo:colect_23_ip_triangle}), 
calculation of edge-triangle ICs (Algorithm \ref{algo:collect_23_ip_edges}), oredering of ICs (Algorithm \ref{algo:order_23}).
The intersection corners appended to the list $I$ as they are computed however their order on the polygon boundary is 
defined by the \emph{connection tables} $F_g(:)$ and $F_p(:)$. Every side of the polygon that lies on \nface $x\in M_2\cup M_3$ is followed by an IC given by $F_g[x]$ 
and every IC $p$ is followed by the side that lies on $F_p[p]\in M_2\cup M_3$ (see \noteJB{possible figure}). The vertices of the polygon are numbered in the same 
way as is the order and orientation of the sides of $S_2$, that is counterclockwise around the interiror with normal pointing to us.

\begin{algorithm}
  \caption{2d-3d intersection, points on triangle boundary}
  \label{algo:colect_23_ip_triangle}

@@

\begin{algorithm}  
  \caption{2d-3d intersection, points in triangle interior}
  \label{algo:collect_23_ip_edges}

@@

  \lFor{edge $e$ of $S_3$}{$L[e]$ = intersection($e, S_2$)}  \nllabel{line:12edges}
  \For{edge $e$ of $S_3$ with non-empty $L[e]$}{
    $p=L[e]$\;
    \uIf{ $p$ is inside $e$}{ 
        \nllabel{line:edge_faces}        
        $(f_0,\ f_1)$ = \edgefaces($e$)
    } 
    \Else($p$ at the vertex $v$ of $S_3$){       
            $(f_0,\ f_1)$ = \vertexfaces($v$,$L$)   \tcp*{Algorithm \ref{algo:vertex_faces}}
            \nllabel{line:mark_edges}
    }    
    \uIf{ $p$ is on boundary of $S_2$ }{
         $p$ lies on edge or at vertex $x \in M_3$\;
         $q=F_g[x]$\tcp*{$q$ is already computed $p$}
         \lIf{$F_p[q]=x$}{$F_p[q]=f_1$}
         \lElse{$F_g[f_0]=q$}
         $F_g[x]=-1$\tcp*{remove the backlink}
    }
    \Else{
         append $p$ to $I$\;
         $F_g[f_0]=p,\ F_p[p]=f_1$\tcp*{overwrite the backlink}
    }   
  }  
\end{algorithm}

Algorithm \ref{algo:collect_23_ip_edges} uses the line-triangle intersection algorithm for the edges of $S_3$ (line \ref{line:12edges}).
First, the intersection $L[e]$ is evaluated for every edge $e$. The loop produces ICs in the interior of $S_2$ and possibly those ICs with special position on vertex or edge of $S_3$ already computed 
in Algorithm \ref{algo:colect_23_ip_triangle}. Then we pass through once again skipping edges with none or degenerate $L[e]$.
For every IC $p=L[e]$ we first get its (generalized) faces that are before and after the IC on IP. For an IC inside the edge $e$
the function \emph{edge faces}, returns its adjacent faces $f_0$, $f_1$ (see Fig. \ref{fig:edge_faces}). 
Their order is given by the sign permuted inner products in 1d-2d intersection. The order of faces match 
the order of sides of IP if the sign is negative. If the sign is positive the function \emph{edge faces} returns face pair $(f_1, f_0)$.
\begin{figure}[!htb]
  \begin{center}        
    \includegraphics[width=0.3\textwidth]{\figpath edge_faces.pdf}
  \end{center}
  \caption{Order of faces adjacent to the oriented edge $e$ pointing up.}
  \label{fig:edge_faces}
\end{figure}

The function \emph{vertex faces} (Algorithm \ref{algo:vertex_faces}, described later)
returns a pair of generalized faces (face or edge) possibly adjacent to the IC $L[e]$ at the vertex $v$ of $S_3$.
On line \ref{line:mark_edges} we mark all edges coincident with the vertex, since if they have any other intersection corner they are degenerate and 
not processed anyway.

@@

  \uElseIf{ $s$ have 1 degenerate edge $e$}{\nllabel{line:one_degenerated}
    $f$ is face opposite to $e$
    \uIf{other two edges $e_a$, $e_b$ have different sign}{
        $z=\edgefaces(e_a)$\;
        replace $g\in z$, $g\ne f$ with $e$
        \return $z$
    }\lElse{
        append IC of $v$ to $J$
        \return anything\;
    }
  }

  \uElseIf{ $s$ have edge $e$ with sign oposite to other two}{ \nllabel{line:different_sign}
    \return edge faces($e$)\;
  }

  \Else($s$ have all signs same){ \nllabel{line:same_signs}
    append IC of $v$ to $J$
    \return anything\;
  }  
\end{algorithm}

\subsubsection{Vertex Faces Algorithm}
This function gets an IC $p$ at vertex $v$ of $S_3$ as a parameter. The IC is special vertex case of non-degenerate edge-triangle intersection. 
The function returns a pair of generalized faces of $S_3$ preceeding and succeeding $p$ on the polygons boundary in the case that $p$ is at interior of $S_2$.
The basic idea is to use the signs of ICs of the three edges coincident with $v$. 
Possible cases are:
\begin{itemize}
 \item {\bf All ICs have the same sign.} (line \ref{line:same_signs}) We return any pair of faces.  $S_2$ is touching $S_3$ 
 at the vertex $v$, the polygon degenerates into the single IC $p$, no connection information from table $F$ is necessary.
 \item {\bf Single IC has the opposite sign to the other two.} (line \ref{line:different_sign}) Let $e$ be the edge of the single IC with the different sign. 
 The plane of $S_2$ separates $e$ from the other two edges so it goes through the faces adjacent to $e$. The order is detemined by the function \emph{edge faces}.
 \item {\bf Single degenerated IC.} (line \ref{line:one_degenerated}) Let us denote $e$ the edge with degenerated IC and $f$ the face between the other two edges.
 The other two (non-degenerates) edges may have either the opposite sign (the plane is cutting $S_3$) or the same sign (the plane is touching $S_3$ at the edge $e$).
 In the first case, the call of edge faces for $e$ returns $(f_x,f)$ or $(f,f_x)$, then the vertex faces function returns $(e, f)$ or $(f, e)$, respectively.

 The edge $e$ lies in the the singel edge
 \item {\bf Two degenerated ICs.} (line \ref{line:two_degenerated}) A face of $S_3$ lies in the plane of $S_2$, single edge $e$ have non-degenerate IC.
 We treat the two degenerate edges as special case of faces adjacent to $e$ and return them sorted like the faces given by edge faces of edge $e$.
\end{itemize}

% Either all signs are equal ($S_2$ touching $S_3$) wee , or
% a single edge have the sign oposite to other twotwo edges have a sign oposite then the thi. of According to the 
% 
% 
% and need to insert them into $I$ to correct place. To this end we exploit additional information about coincidence of points
% with faces of $S_3$. Every point $p$ from the first part of the algorithm \notePE{I would use proper references instead, I got confused, what 'part of algorithm' is..} that is not a vertex of $S_2$ belongs to a face $f$ of $S_3$; we know from the order of the points
% whether the face comes before or after the point in the intersection polygon; in the first case we set $F[0,f] =p$, in the second case we set $F[1,f]=p$. 
% Every point $p$ from the second part of the algorithm lies on an edge $e$ which is between two faces $f_0$, $f_1$. The correct order of the faces on the 
% intersection polygon is determined by the sign of the edge-triangle intersection. So, having the faces in the correct order we set $F[0,f_0]=p$ and $F[1, f_1]=p$.

Finally in the third part (Algorithm \ref{algo:order_23}), the table $F$ allows us to modify array of successors $P$ and get $I$ in the correct order as the list $K$.

\begin{algorithm}
  \caption{2d-3d intersection, order the points}  
  \label{algo:order_23}

  \SetKw{To}{to}
  \DontPrintSemicolon
  \KwIn{all points in $I$}
  \KwOut{polygon in $K$ in correct order}

  \lIf{$\abs{I} < 3$}{ return $J$}
  \lElse{ return $I$ sorted according to connectivity in $F_g$ and $F_p$}
\end{algorithm}
\noteJB{Write correct algo with deduplication of vertices.}

 List $in[f]$ contains index of the intersection corner that follows after $f$ on the boundary of traced polygon, 
 similarly $out[f]$ stores index of the intersection corner that preeceeds the face $f$.

Possible cases for processing $L$:
\begin{enumerate}
 \item Regular case, $L$ consists of two intersections $p$, $q$ sorted by orientation of $s$, laying inside of $s$.

 If $p$ is on the edge $e$ of $S_3$ compute sign of intersection($e$, $S_2$), sort the faces $f_0$, $f_1$ coincident with $e$
 and set $in[f_0]$ to index of $p$ in $L$. Similarly if $q$ is on the edge, set $out[f_1]$ to index of $q$ in $L$.

 If $p$ is in vertex $v$ of $S_3$, for every face $f$ coincident with $v$ set
 in[$f$] to index of $p$ unless there is some index already set. So, we do not over ride
 entries comming from the edge intersections. Similarly set out[$f$] if $q$ is in vertex of $S_3$.

 If $p$ is on face $f_0$ of $S_3$, set $in[f_0]$ to index of $p$. Similarly, if $q$ is on face $f_1$ of $S_3$, set $out[f_1]$ to index of $q$.
 laying on faces $f_p$, $f_q$ of $S_3$.

 \item $L$ consists of a single intersection corner $p$ (touching $S_3$)

 If $p$ is on edge, compute sign of intersection($e$, $S_2$), sort the faces $f_0$, $f_1$,
 set $in[f_0]$ and $out[f_1]$ to index of $p$.

 If $p$
\end{enumerate}

How tracing works.
\begin{itemize}
 \item If there are no intersections in vertex of $S_3$. 
\end{itemize}

intersection polygon
are found as intersection corners of either triangle side and tetrahedron or tetrahedron edge and triangle.
Therefore we use both algorithms above for 1D-3D and 1D-2D, respectively. Data are again efficiently passed
to lower dimensional problems, so

The array of intersection corners is generally not sorted. We use two so called \emph{tracing} algorithms and we
intend to orient the edges of the polygon in the same direction as the triangle is oriented.
If one of the intersection corner is pathologic, a general convex hull method is applied using the Monotone 
chain\footnote{Wikibooks, [online 2016-03-01], 
      \url{http://en.wikibooks.org/wiki/Algorithm_Implementation/Geometry/Convex_hull/Monotone_chain}} 
algorithm. The points are sorted using only their barycentric coordinates.

An optimized algorithm has been suggested for non-pathologic cases. At this moment all the collected topology 
data come into play. The algorithm takes advantage by using only the data already computed and also lowers
the complexity to $O(N)$, compared with the Monotone chain complexity $O(N\log N)$ 
($N$ being number of intersection corners).

\subsection{Tracking boundary of the intersection polygon}
\label{sec:front advancing}
\notePE{It seems to me, that we can really describe the prolongation in general, for 1D and 2D, 
refering to them as components. I tried to do so..}

\notePE{We need proper definitions of terms we use: \\
% \emph{component} is a set of connected elements of the same lower dimension (1D, 2D)\\
% \emph{bulk} is a set of 3D elements\\
% or slave, master...
\emph{candidates pair}  is a pair of a component element and a bulk element, that might intersect
each other (due to intersection of their bounding boxes or prolongation result)
\emph{pathologic, special, degenerate case ??}
}

\notePE{Do you want to use American 'neighbor' or all other English 'neighbour'? (I prefer non-american.. to be unified at the end..)}

>>>>

\subsection{Intersection Triangle-Tetrahedron (2D-3D)}
The intersection of a triangle $S_2$ and a tetrahedron $S_3$ is an $n$-side intersection polygon (IP), $n\le 7$. The sides of the polygon
lie either on sides of $S_2$ or on faces of $S_3$. Thus each vertex (IC) of the polygon
can arise either from side-face intersection, or from edge-triangle intersection, or be a vertex of $S_2$.
To get all ICs, we have to compute at most 12 side-face intersections and at most 6 edge-triangle intersections. However,
to this end we only need to compute 9 \plucker coordinates (3 sides, 6 edges) and 18 permuted inner produts, one for every side-edge pair.
Computation of IP consists of three stages: calculation of side-tetrahedron ICs (Section \ref{sec:sides}), 
calculation of edge-triangle ICs (Section~\ref{sec:edges}), oredering of ICs (Section~\ref{sec:ordering}).
The intersection corners are appended to the list $I$ as they are computed however their order on the polygon boundary is 
defined by the \emph{connection tables} $F_g(:)$ and $F_p(:)$. Every side of the polygon that lies on \nface $x\in M_2\cup M_3$ is followed by an IC given by $F_g[x]$ 
and every IC $p$ is followed by the side that lies on $F_p[p]\in M_2\cup M_3$ (see \noteJB{possible figure}). The vertices of the polygon are oredered in the same 
way as is the order and orientation of the sides of $S_2$, that is counterclockwise around the interior with normal pointing to us.
There are special cases when the IP is degenerated to a line or a point. If some IC may possibly be part of the such degenerated IP we put it into the list $J$.

\subsubsection{Intersections on sides of $S_2$}
\label{sec:sides}

\begin{algorithm}
  \caption{2d-3d intersection, ICs on sides of $S_2$}
  \label{algo:colect_23_ip_triangle}

@@

\subsubsection{Intersections on edges of $S_3$}
\label{sec:edges}
\begin{algorithm}  
  \caption{2d-3d intersection, ICs on edges of $S_3$}
  \label{algo:collect_23_ip_edges}

@@

  \lFor{edge $e$ of $S_3$}{$L[e]$ = intersection($e, S_2$)}  \nllabel{line:12edges}
  \For{edge $e$ of $S_3$ with non-empty $L[e]$}{
    $p=L[e]$\;
    \uIf{ $p$ is inside $e$}{ 
        \nllabel{line:edge_faces}        
        $(f_0,\ f_1)$ = \edgefaces($e$)
    } 
    \Else($p$ at the vertex $v$ of $S_3$){       
            $(f_0,\ f_1)$ = \vertexfaces($v$,$L$)   \tcp*{Algorithm \ref{algo:vertex_faces}}
            \nllabel{line:mark_edges}
    }    
    \uIf{ $p$ is on boundary of $S_2$ }{
         $p$ lies on edge or at vertex $x \in M_3$\;
         $q=F_g[x]$\tcp*{$q$ is already computed $p$}
         \lIf{$F_p[q]=x$}{$F_p[q]=f_1$} \nllabel{line:successor}
         \lElse{$F_g[f_0]=q$}  \nllabel{line:predecessor}
         $F_g[x]=-1$\tcp*{remove the backlink}
    }
    \Else{
         append $p$ to $I$\; 
         $F_g[f_0]=p,\ F_p[p]=f_1$\tcp*{overwrite the backlink} \nllabel{line:insideS2}
    }   
  }  
\end{algorithm}

\begin{figure}%[!htb]
  \begin{center}        
    \includegraphics[width=0.3\textwidth]{\figpath edge_faces.pdf}
  \end{center}
  \caption{Order of faces adjacent to the oriented edge $e$ pointing up.}
  \label{fig:edge_faces}
\end{figure}

Algorithm \ref{algo:collect_23_ip_edges} uses the line-triangle intersection algorithm for the edges of $S_3$ (line \ref{line:12edges}).
First, the intersection $L[e]$ is evaluated for every edge $e$. The loop produces ICs in the interior of $S_2$ and possibly those ICs with special position on vertex or edge of $S_3$ already computed 
in Algorithm \ref{algo:colect_23_ip_triangle}. Then we pass through once again skipping edges with none or degenerate $L[e]$.
For every IC $p=L[e]$ we first get its (generalized) faces that would appear before and after the IC on IP assuming that $p$ is inside $S_2$.

For an IC inside the edge $e$
the function \emph{edge faces}, returns its adjacent faces $f_0$, $f_1$ (see Fig. \ref{fig:edge_faces}). 
Their order is given by the sign permuted inner products in 1d-2d intersection. The order of faces match 
the order of sides of IP if the sign is negative. If the sign is positive the function \emph{edge faces} returns face pair $(f_1, f_0)$.
If the IC is at the vertex $v$ of $S_3$, the function \emph{vertex faces} described later (Algorithm \ref{algo:vertex_faces}) is used.
It returns a pair of generalized faces (face or edge) adjacent to the IC $L[e]$ at the vertex $v$ of $S_3$. 
If IC $p$ is inside $S_2$, we add it into $I$ and set the connection tables (line \ref{line:insideS2}). However,
if $p$ is on boundary of $S_2$ it is already in $I$. Denoting $x$ the \nface of $S_3$ the IC lies on, we use the back-link from $x$ to 
IC $q\in I$ and if $q$ points back to $x$ we set the new successor $f_1$ to it (line \ref{line:successor}). 
Otherwise $f_0$ is used as the predecessor of $q$ (line \ref{line:predecessor}).

\subsubsection{Vertex Faces Algorithm}

@@

  \uElseIf{ $s$ have 1 degenerate edge $e$}{\nllabel{line:one_degenerated}
    $f$ is face opposite to $e$\;
    \uIf{other two edges $e_a$, $e_b$ have different sign}{
        $z=\edgefaces(e_a)$\;
        replace $g\in z$, $g\ne f$ with $e$, \return $z$
    }\lElse{
        append IC of $v$ to $J$, \return anything
    }
  }  
  \uElseIf{ $s$ have edge $e$ with sign oposite to other two}{ \nllabel{line:different_sign}
    \return edge faces($e$)\;
  }

  \Else($s$ have all signs same){ \nllabel{line:same_signs}
    append IC of $v$ to $J$, \return anything\;
  }  
\end{algorithm}

This function gets as a parameter IC $p$ at the vertex $v$ of $S_3$ which is a special case of anon-degenerate edge-triangle intersection. 
We use signs and degenerate indicator of ICs of the three edges coincident with $v$ to return generalized faces of $S_3$ 
preceding and succeeding $p$ on the polygons boundary assuming $p$ is at interior of $S_2$.
Possible cases are:
\begin{itemize}
 \item {\bf Single degenerated IC} (line \ref{line:one_degenerated}){\bf .} Let us denote $e$ the edge with degenerated IC 
 and $f$ the face between the other two edges.
 The other two (non-degenerates) edges may have either the opposite sign (the plane is cutting $S_3$) 
 or the same sign (the plane is touching $S_3$ at the edge $e$).
 In the first case, the call of edge faces for $e$ returns $(f_x,f)$ or $(f,f_x)$, then the vertex faces function returns $(e, f)$ or $(f, e)$, respectively.
 In the second case, we just return anything since there are at most two ICs in IP so the connection tables are not necessary. However, we add 
 current IC into $J$.

 \item {\bf Two degenerated ICs} (line \ref{line:two_degenerated}){\bf .} A face of $S_3$ lies in the plane of $S_2$. 
 Let $e$ be the single non-degenerate edge. We treat the two degenerate edges as faces adjacent to $e$ 
 and return them sorted like the faces given by edge faces of edge $e$.

 \item {\bf Single IC has the opposite sign to the other two} (line \ref{line:different_sign}){\bf .} 
 Let $e$ be the edge of the single IC with the different sign. 
 The plane of $S_2$ separates $e$ from the other two edges so it goes through the faces adjacent to $e$. 
 The order is determined by the function \emph{edge faces} called for the edge $e$.

 \item {\bf All ICs have the same sign} (line \ref{line:same_signs}){\bf .} Since $S_2$ is touching $S_3$ 
 at the vertex $v$, the polygon degenerates into point and thus no connection information is necessary. We add IS to $J$ and return any pair of faces.
\end{itemize}

% Either all signs are equal ($S_2$ touching $S_3$) wee , or
% a single edge have the sign oposite to other twotwo edges have a sign oposite then the thi. of According to the 
% 
% 
% and need to insert them into $I$ to correct place. To this end we exploit additional information about coincidence of points
% with faces of $S_3$. Every point $p$ from the first part of the algorithm \notePE{I would use proper references instead, I got confused, what 'part of algorithm' is..} that is not a vertex of $S_2$ belongs to a face $f$ of $S_3$; we know from the order of the points
% whether the face comes before or after the point in the intersection polygon; in the first case we set $F[0,f] =p$, in the second case we set $F[1,f]=p$. 
% Every point $p$ from the second part of the algorithm lies on an edge $e$ which is between two faces $f_0$, $f_1$. The correct order of the faces on the 
% intersection polygon is determined by the sign of the edge-triangle intersection. So, having the faces in the correct order we set $F[0,f_0]=p$ and $F[1, f_1]=p$.
\subsubsection{Ordering of intersections}
\label{sec:ordering}
The final stage of the 2d-3d intersection is ordering of ICs. If there are less then 3 ICs in the list $I$ we merge lists $I$ and $J$ remove duplicities
and get at most 2 distinguish ICs forming a degenerate IP. Otherwise, we start with first IC $p$ in $I$, put it into final list $K$, 
get its successor \nface $x = F_p(p)$ and its successor IC $q=F_q(x)$. Unless $x$ is vertex of $S_2$, we put $q$ into $K$ and repeat the process until 
we pass all ICs in $I$.
